---
title: "US Maps"
---

```{r}
#| message: false
#| warning: false
#| echo: false
#| output: false

library(tidyverse)
library(maps)
library(viridis)
library(mdsr)

library(readr)

min_wage <- read_csv("Minimum Wage Data.csv")
us_states <- map_data("state")

min_wage|>
  distinct(Year, State)|>
  count(Year)
```

```{r}
#| message: false
#| warning: false
#| echo: false


min_wage|>
  select(Year, State, State.Minimum.Wage) |>
  filter(Year %in% c(2017, 2018, 2015, 2016, 2012, 2020))|>

  
  right_join(us_states, by= c("State"= "region"))|>
  ggplot(mapping=aes(x=long, y=lat,
                     group=group))+
  geom_polygon(aes(fill=State.Minimum.Wage), color="black")+
  labs(fill= "Population Percent Change", title = "Population Change Across the U.S. States",
       subtitle = "Visualizing Percent Change in Population by State")+
  coord_map()+
  theme_gray()+
  scale_fill_gradient()+
  facet_wrap(~Year) 
```

### Exploring Minimum Wage Trends Across the U.S.

This analysis dives into the changing landscape of minimum wages across the United States over selected years, highlighting the shifts and trends from 2012 to 2020.

Through these map, we're not just looking at numbers; we're seeing the impact of policy changes, economic shifts, and the ongoing debate around fair wages and living standards.
